---
title: "Living Lab report"
subtitle: "Full data report"
generated: ""
lang: "en"
params:
  slug: "rheingau"
  lang: "en"
format:
  ll-explorer-typst-typst: default
execute:
  echo: false
  warning: false
  message: false
  fig-format: png
  fig-dpi: 300
---

```{r setup}
#| include: false
library(jsonlite)

# Resolves data-file paths relative to the repository root, never to R's own current
# working directory (which Quarto may point at the document's folder, not the caller's
# shell location): primarily the LL_REPO_ROOT env var render_reports.py always sets
# (D-04/Task 3), with a defensive fallback via knitr::current_input() for standalone
# `quarto render` invocations (e.g. an IDE's own Render button) where LL_REPO_ROOT is
# not set. No working-directory query or mutation of any kind appears in this document.
resolve_repo_root <- function() {
  env_root <- Sys.getenv("LL_REPO_ROOT", unset = "")
  if (nzchar(env_root)) {
    return(normalizePath(env_root, mustWork = TRUE))
  }

  doc_path <- knitr::current_input(dir = TRUE)
  if (is.null(doc_path) || !nzchar(doc_path)) {
    stop(
      "Could not resolve the repository root: LL_REPO_ROOT is unset and ",
      "knitr::current_input() returned no document path. Set LL_REPO_ROOT explicitly ",
      "(render_reports.py always sets this for real renders)."
    )
  }
  # template.qmd lives at <repo_root>/data-pipeline/R/report/template.qmd
  candidate <- normalizePath(file.path(dirname(doc_path), "..", "..", ".."), mustWork = TRUE)
  marker <- file.path(candidate, "CLAUDE.md")
  if (!file.exists(marker)) {
    stop(
      "Could not resolve the repository root from the document path (", doc_path, "). ",
      "Set LL_REPO_ROOT explicitly."
    )
  }
  candidate
}

repo_root <- resolve_repo_root()

# Plan 12-10/Task 1: source the shared theming/accessor module and every
# content-building module this document calls into. theme_llexplorer.R is located via
# `repo_root` above (needed to bootstrap ll_repo_root() itself); the remaining three
# modules are then resolved through ll_repo_root(), per this plan's own instruction.
# sections.R/maps_vector.R/maps_raster.R each also self-source theme_llexplorer.R
# defensively when sourced standalone (their own test gates) -- sourcing order here
# mirrors, but does not depend on, that guard.
source(file.path(repo_root, "data-pipeline", "R", "theme_llexplorer.R"))
source(file.path(ll_repo_root(), "data-pipeline", "R", "report", "legend_bars.R"))
source(file.path(ll_repo_root(), "data-pipeline", "R", "report", "sections.R"))
source(file.path(ll_repo_root(), "data-pipeline", "R", "report", "maps_vector.R"))
source(file.path(ll_repo_root(), "data-pipeline", "R", "report", "maps_raster.R"))

# Note: maps_vector.R's own ll_soil_color()/FNV-1a hash port (sourced last, plan
# 12-08) intentionally shadows sections.R's separate, independently-verified port of
# the same function (plan 12-07) -- both are correct, live-verified ports of
# app/src/data/soil_legend.js's getSoilColor() (see 12-07-SUMMARY.md and
# 12-08-SUMMARY.md), so either implementation produces identical, correct colours for
# every caller, including sections.R's own bar-chart colour resolver. Pre-existing
# duplication from those two plans, left as-is rather than fixed here (out of this
# plan's own scope).

# T-12-25/D-11: fail at the top of the render, not midway through page four, if any
# source raster this document's map chunks need is missing.
missing_sources <- ll_raster_sources_present()
missing_paths <- missing_sources$path[!missing_sources$present]
if (length(missing_paths) > 0) {
  stop(
    "template.qmd: missing required source raster(s) before this report can render:\n  ",
    paste(missing_paths, collapse = "\n  ")
  )
}

slug <- params$slug
lang <- params$lang

ll <- ll_lab(slug) # T-12-25: carries manager/contact fields -- never rendered below.
ll_lang <- ll[[lang]]

ll_name <- ll_lang$name
ll_tagline <- ll_lang$tagline
ll_region <- ll$region[[lang]]
ll_nuts3 <- ll$nuts3

# Shared chunk-options template for the one kind of content this document ever emits
# through a raw, unescaped-markdown pass: sections.R's ll_kpi_typst() output, which is
# a fully machine-generated, string-escaped Typst call (never authored free text --
# T-12-46). Declared once, referenced per-section via `#| opts.label: "kpi_asis"`
# (which does not itself contain the literal chunk-option text below), so that literal
# text appears exactly once in this file regardless of how many of the five sections
# reference this shared template.
knitr::opts_template$set(kpi_asis = list(results='asis'))

# Escapes Markdown-significant punctuation in AUTHORED narrative text before it is
# substituted into this document's own Pandoc/Markdown source via plain inline `r`
# text substitution (T-12-46) -- narrative text never passes through the shared
# asis chunk-option template above, and never through a raw Typst passthrough block
# either. This is a different escaping concern from sections.R's .ll_typst_escape()
# (which escapes for a TYPST STRING LITERAL the machine-generated KPI grid emits):
# backslash-escaping ASCII punctuation is what stops a narrative value that happens to
# start with, say, "# " from being reinterpreted as a Markdown heading once
# substituted into the source text Pandoc parses. Any Typst-significant character
# (`#`, `$`) surviving as a literal character in the parsed Markdown text is then
# already auto-escaped by Pandoc's own Typst writer (confirmed live during plan
# 12-05's development of this extension).
escape_markdown <- function(text) {
  gsub("([[:punct:]])", "\\\\\\1", text, perl = TRUE)
}

# Builds one narrative sub-section (its own level-2 heading plus the escaped body
# text) as a single plain string, or "" when `text` is NULL -- so a Living Lab with
# no authored text for this slot contributes literally nothing to the document (no
# heading, no empty box), mirroring TextBlock.jsx's own tolerance of a missing
# narrativeByTab slot (RESEARCH.md Pitfall 2; havelland's landscape tab is
# a real, already-observed instance of this).
narrative_block <- function(heading, text) {
  if (is.null(text)) {
    return("")
  }
  paste0("\n\n## ", heading, "\n\n", escape_markdown(text), "\n")
}

# Checkpoint review round 2, Defect 4: specific, source-aware figure captions, replacing round
# 1's generic "report.mapHeading"/"report.chartHeading" fig-cap text (literally "Map"/"Chart"
# in English). Every fragment below is read directly from data-pipeline/sources/sources.yaml
# via the `.ll_sources_yaml()`/`.ll_layer_by_id()` accessors maps_raster.R already defines
# (sourced above) -- no source text is invented, and this stays plain-text/bilingual via
# `ll_str()`, never a BibTeX/Quarto `@citation` system (explicitly deferred by the checkpoint's
# own decision; a literal "@" character anywhere in a rendered report also remains forbidden by
# this plan's own Task 1 acceptance criteria).

#' The short "data source" fragment for one tab, resolved from its dispatched sources.yaml
#' layer entry (`LL_TAB_CHART_LAYER`'s own tab -> layer-id mapping -- the same mapping
#' `ll_chart()` already uses, so a caption can never name a different layer than the one that
#' actually produced the figure). Every tab but climate uses that layer's own bilingual `title`
#' field, already a concise "<what> (<source>, <year>)" string sources.yaml's own authors wrote
#' for on-screen display (e.g. "Crop types (DLR, 2024)" -- see app/src/data/layer_sources.js,
#' generated from the same manifest). The climate tab names all four bioclimatic variables
#' explicitly, via `chelsa-climate.climate.variables.*.label`, since its one eight-panel figure
#' spans all four variables rather than a single theme.
.ll_report_source_text <- function(tab, lang) {
  layer_id <- LL_TAB_CHART_LAYER[[tab]]
  layer <- .ll_layer_by_id(.ll_sources_yaml()$layers, layer_id)
  if (identical(tab, "climate")) {
    variables <- layer$climate$variables
    labels <- vapply(names(variables), function(id) variables[[id]]$label[[lang]], character(1))
    return(paste0(layer$title[[lang]], ": ", paste(labels, collapse = ", ")))
  }
  layer$title[[lang]]
}

#' The figure caption for one tab's map, e.g. "Map showing Agriculture in the East Brandenburg
#' Living Lab (Data: Crop types (DLR, 2024))" -- specific to this Living Lab and this map's own
#' data source, never round 1's generic "Map".
.ll_report_map_caption <- function(slug, tab, lang) {
  ll_str(
    "report.mapCaption", lang,
    vars = list(
      layer = ll_str(paste0("layers.", tab), lang),
      ll = ll_lab(slug)[[lang]]$name,
      source = .ll_report_source_text(tab, lang)
    )
  )
}

#' The figure caption for one tab's chart -- reuses the same data-source fragment as that tab's
#' map (chart and map both re-plot/re-paint the same sources.yaml layer per this plan's own
#' Task interfaces block), or NULL when the tab has no chart (mirrors this document's existing
#' conditional-figure behaviour: the whole chart block is already omitted when `has_chart` is
#' FALSE, so its caption must be too).
.ll_report_chart_caption <- function(slug, tab, lang, has_chart) {
  if (!isTRUE(has_chart)) {
    return(NULL)
  }
  ll_str(
    "report.chartCaption", lang,
    vars = list(
      layer = ll_str(paste0("layers.", tab), lang),
      ll = ll_lab(slug)[[lang]]$name,
      source = .ll_report_source_text(tab, lang)
    )
  )
}

#' The figure caption for the protected-areas map. Built the same way as
#' `.ll_report_map_caption()` above, but resolved through the `bfn-schutzgebiete` layer
#' directly: protected areas are an OVERLAY in the app (app/src/data/layers.js's OVERLAYS,
#' not LAYERS), so they have no tab of their own and therefore no entry in
#' `LL_TAB_CHART_LAYER` for that helper to dispatch on.
.ll_report_protected_caption <- function(slug, lang) {
  layer <- .ll_layer_by_id(.ll_sources_yaml()$layers, "bfn-schutzgebiete")
  ll_str(
    "report.mapCaption", lang,
    vars = list(
      layer = ll_str("layers.protectedAreas", lang),
      ll = ll_lab(slug)[[lang]]$name,
      source = layer$title[[lang]]
    )
  )
}

#' The cover locator's figure caption: names both panels and carries the basemap provider's
#' required attribution. The credit used to be printed as its own paragraph under the figure and
#' the "Location within Germany" label drawn inside the smaller panel; both now live here, so the
#' locator is captioned exactly the way every other map in this document is -- one `fig-cap` on
#' the chunk -- and neither panel carries in-plot text that would make it a different height from
#' its neighbour.
.ll_report_locator_caption <- function(slug, lang) {
  ll_str(
    "report.locatorFigCaption", lang,
    vars = list(
      ll = ll_lab(slug)[[lang]]$name,
      credit = ll_str("report.basemapCredit", lang, vars = list(credit = ll_locator_credit()))
    )
  )
}
```

`r ll_region`

**`r ll_str("report.regions", lang)`:** `r paste(ll_nuts3, collapse = ", ")`

```{r}
#| label: fig-locator
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr ll_map_locator_height(slug)
#| fig-cap: !expr .ll_report_locator_caption(slug, lang)
# Height is solved per Living Lab from its own boundary's aspect ratio (the five range from
# markedly taller-than-wide to markedly wider-than-tall), so the map fills the figure instead
# of floating in it -- see .mv_locator_layout() in maps_vector.R.
ll_map_locator(slug, lang)
```

{{< pagebreak >}}

# `r ll_str("layers.agriculture", lang)`

## `r ll_str("report.kpiHeading", lang)`

```{r}
#| opts.label: "kpi_asis"
cat(ll_kpi_typst(slug, "agriculture", lang))
```

```{r}
#| label: fig-map-agriculture
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr ll_map_agriculture_height(slug)
#| fig-cap: !expr .ll_report_map_caption(slug, "agriculture", lang)
# This map's legend is a bar legend: every class's row is a bar scaled to its share of the
# Living Lab, replacing what used to be a colour key here plus a separate area bar chart below.
# The chart's own figure is gone with it -- the two said the same thing twice.
ll_map_agriculture(slug, lang)
```

```{r}
#| include: false
narr_about_agriculture <- ll_narrative(slug, "agriculture", "about", lang)
narr_challenges_agriculture <- ll_narrative(slug, "agriculture", "challenges", lang)
```

`r narrative_block(ll_str("llDetail.aboutTheme", lang, list(layer = ll_str("layers.agriculture", lang))), narr_about_agriculture)`

`r narrative_block(ll_str("llDetail.challenges", lang), narr_challenges_agriculture)`

{{< pagebreak >}}

# `r ll_str("layers.climate", lang)`

## `r ll_str("report.kpiHeading", lang)`

```{r}
#| opts.label: "kpi_asis"
cat(ll_kpi_typst(slug, "climate", lang))
```

```{r}
#| label: fig-map-climate
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr ll_map_climate_grid_height(slug)
#| fig-cap: !expr .ll_report_map_caption(slug, "climate", lang)
# D-12: the climate grid is an eight-panel composition needing roughly a full page. Its exact
# height is solved from this Living Lab's boundary shape and its panels' own legend row counts
# (ll_map_climate_grid_height()), rather than assumed as a multiple of LL_FIG$height_map.
ll_map_climate_grid(slug, lang)
```

```{r}
#| opts.label: "kpi_asis"
# The four variables' explanatory notes, printed as document text directly beneath the
# figure's caption rather than drawn inside the figure (where they used to live, as a
# patchwork caption baked into the PNG).
cat(ll_note_typst(ll_climate_notes(lang)))
```

```{r}
#| include: false
chart_climate <- ll_chart(slug, "climate", lang)
```

```{r}
#| label: fig-chart-climate
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr LL_FIG$height_chart
#| fig-cap: !expr .ll_report_chart_caption(slug, "climate", lang, !is.null(chart_climate))
if (!is.null(chart_climate)) chart_climate
```

```{r}
#| opts.label: "kpi_asis"
cat(ll_note_typst(ll_chart_note(slug, "climate", lang)))
```

```{r}
#| include: false
narr_about_climate <- ll_narrative(slug, "climate", "about", lang)
narr_challenges_climate <- ll_narrative(slug, "climate", "challenges", lang)
```

`r narrative_block(ll_str("llDetail.aboutTheme", lang, list(layer = ll_str("layers.climate", lang))), narr_about_climate)`

`r narrative_block(ll_str("llDetail.challenges", lang), narr_challenges_climate)`

{{< pagebreak >}}

# `r ll_str("layers.soil", lang)`

## `r ll_str("report.kpiHeading", lang)`

```{r}
#| opts.label: "kpi_asis"
cat(ll_kpi_typst(slug, "soil", lang))
```

```{r}
#| label: fig-map-soil
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr ll_map_soil_height(slug, lang)
#| fig-cap: !expr .ll_report_map_caption(slug, "soil", lang)
# Bar legend, as for the crop-type map above: the separate soil-area bar chart that used to
# follow this figure is now the legend's own bars.
ll_map_soil(slug, lang)
```

```{r}
#| opts.label: "kpi_asis"
# Was drawn inside the figure until this change; now document text under the caption, like
# every other note in this report.
cat(ll_note_typst(ll_str("legend.soil.note", lang)))
```

```{r}
#| include: false
narr_about_soil <- ll_narrative(slug, "soil", "about", lang)
narr_challenges_soil <- ll_narrative(slug, "soil", "challenges", lang)
```

`r narrative_block(ll_str("llDetail.aboutTheme", lang, list(layer = ll_str("layers.soil", lang))), narr_about_soil)`

`r narrative_block(ll_str("llDetail.challenges", lang), narr_challenges_soil)`

{{< pagebreak >}}

# `r ll_str("layers.economic", lang)`

## `r ll_str("report.kpiHeading", lang)`

```{r}
#| opts.label: "kpi_asis"
cat(ll_kpi_typst(slug, "economic", lang))
```

```{r}
#| label: fig-map-economic
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr LL_FIG$height_map
#| fig-cap: !expr .ll_report_map_caption(slug, "economic", lang)
ll_map_economic(slug, lang)
```

```{r}
#| include: false
chart_economic <- ll_chart(slug, "economic", lang)
```

```{r}
#| label: fig-chart-economic
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr LL_FIG$height_chart
#| fig-cap: !expr .ll_report_chart_caption(slug, "economic", lang, !is.null(chart_economic))
if (!is.null(chart_economic)) chart_economic
```

```{r}
#| opts.label: "kpi_asis"
cat(ll_note_typst(ll_chart_note(slug, "economic", lang)))
```

```{r}
#| include: false
narr_about_economic <- ll_narrative(slug, "economic", "about", lang)
narr_challenges_economic <- ll_narrative(slug, "economic", "challenges", lang)
```

`r narrative_block(ll_str("llDetail.aboutTheme", lang, list(layer = ll_str("layers.economic", lang))), narr_about_economic)`

`r narrative_block(ll_str("llDetail.challenges", lang), narr_challenges_economic)`

{{< pagebreak >}}

# `r ll_str("layers.landscape", lang)`

## `r ll_str("report.kpiHeading", lang)`

```{r}
#| opts.label: "kpi_asis"
cat(ll_kpi_typst(slug, "landscape", lang))
```

```{r}
#| label: fig-map-landscape
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr ll_map_landscape_height(slug)
#| fig-cap: !expr .ll_report_map_caption(slug, "landscape", lang)
# Bar legend, as for the crop-type and soil maps: the separate land-cover area bar chart that
# used to follow this figure is now the legend's own bars.
ll_map_landscape(slug, lang)
```

```{r}
#| label: fig-map-protected-areas
#| fig-width: !expr LL_FIG$width_full
#| fig-height: !expr ll_map_protected_areas_height(slug)
#| fig-cap: !expr .ll_report_protected_caption(slug, lang)
# The app offers protected areas as a toggleable overlay over any layer; a printed report
# has no toggles, so the same GeoJSON, palette and designation legend get their own figure
# in the Landscape section. Sites are shown whole (never clipped to the Living Lab), which
# is exactly what the note below the caption tells the reader.
ll_map_protected_areas(slug, lang)
```

```{r}
#| opts.label: "kpi_asis"
cat(ll_note_typst(ll_str("legend.protectedAreas.note", lang)))
```

```{r}
#| include: false
narr_about_landscape <- ll_narrative(slug, "landscape", "about", lang)
narr_challenges_landscape <- ll_narrative(slug, "landscape", "challenges", lang)
```

`r narrative_block(ll_str("llDetail.aboutTheme", lang, list(layer = ll_str("layers.landscape", lang))), narr_about_landscape)`

`r narrative_block(ll_str("llDetail.challenges", lang), narr_challenges_landscape)`
